### 2024-03-25 gen/paper/paper.Rmd
paper knits

---
title: Investigating the influence of genre preferences on movie/series rating patterns
  over time
author: "Team 1 - Timo Philipse, Bram Teunissen, Rodrigo Pačeko Rudzājs, Dian de Ridder & Linda van den Boogaart"
output: 
  pdf_document: {}
editor_options:
  markdown:
    wrap: 72
---

# Research motivation

The entertainment industry is highly dynamic. Consumer preferences keep
changing, and new movies and series are produced regularly. These
changes in preferences require the industry to respond accordingly. For
example, showing an ankle in a movie might have been considered
scandalous in the 1950s, but it is now just a minor detail. To make an
impact in such an ever-evolving industry, movie directors need to get
creative. These changes in the industry influence how movies and series
are produced and how consumers react to them. One way to understand
these patterns is by examining the genre of a product.

Genre preferences reflect what consumers have enjoyed watching in the
past, and hence are likely to enjoy in the future. Therefore, genre can
be used as a criterion for selecting productions, based on which
consumers decide whether to watch a particular production or not. An
understanding of genre preferences provides filmmakers with guidance on
which genre to choose to increase the chances of success and to best
convey their intended message. However, the nature of genres is
constantly evolving, and it is important to comprehend how current
consumers respond to the evolution of a genre over time. For instance,
an action movie today might have rapid scene changes, multiple plot
twists, and lengthy fight scenes, but on average, such movies may only
receive moderate ratings at best. In contrast, action movies from two
decades ago had a more gradual plot development and were less
overwhelming, yet they received a higher average rating. When consumers
rely on the genre productions of the past to make their choice, it may
adversely impact their evaluation of a current production in the same
genre.

In order to gain a better understanding of how modern consumers rate
products and services, it is crucial to comprehend how they use their
past preferences as a point of reference. Hence, this research paper
aims to explore whether there exist any discernible patterns in genre
preferences over time, and whether some genres tend to have a greater
impact on average ratings than others.

## Research question

In this paper we aim to answer the following research question:

**How do consumers' production preferences of genre elements change over
time, and how do they impact the average ratings of those same
productions?**

## Relevance

The research results shed light on how consumers use dynamic genres to
shape their satisfaction with a movie or TV series. The findings suggest
that some genres are preferred in certain years over others. For
example, war movies tend to receive higher ratings during world wars
compared to times of peace. This may imply that consumers use these
productions as references to rate them in context of the sign of the
times. Managers may want to consider adjusting their productions to
better suit the elements in those preferred years.

The study also found that recent fantasy movies tend to receive lower
ratings from consumers, while those produced in the 70s are rated more
favorably. Overall, the results may suggest that managers may face
challenges in meeting previously set expectations and that these
findings provide a reference frame for when those expectations are met.

Additionally, the results also provide insights into how genres are
related to average ratings across another. Currently, every production
has a similar pattern of ingredients. Most movies contain hints of drama
and romance, regardless of whether it is in a distropian sci-fi survival movie or an horror action series. However, the results indicated that these common genres are not significantly related to the average ratings. In contrast some movies and series do contain unique elements do results in
significant relationships. For example, biographies and animations. This
may imply that consumers may not rely so heavily on past genres that are
common because they gradually change with times without consumers
noticing. (Think of it like a birthday party with your family when you
are 12. Your aunt keeps telling you "How much you have grown!", but you
don't notice any difference) In contrast, the uniqueness of biographies
does sharply stand out. Therefore, managers may want to focus more
strongly on the less common genre features compared to more common ones.

## Research method

In this paper we investigate consumers’ movie and series preferences
using four of IMDb's data sets containing movie data. First, these

Data sets are loaded, cleaned, inspected, merged and cleaned. Next, the
data is analyzed using a regression analysis. This method was chosen due to the continuous nature of the dependent variable and the categorical nature of the independent variables. The reason why a regression was chosen over the alternative of an ANOVA was due to the size and complexity of the analysis. There are 27 genres included as a dependent variable, and are measure over time. To include these features into an ANOVA is far more complex compared to a regression analysis.

This regression analysis regresses all movie and series genre's against the average rating's over a sliding period of 5 years. Additionally, the variables title, movieType and time are used as control variables to improve the interpret ability of the estimates. Additionally, because some of the assumptions were only robustly satisfied, the standard errors were clustered. The analysis resulted in a 20 window output, one for each year window.

A note of attention should be, that despite the richness of the data, the interpretation of the results remain descriptive at bests. This is because the average movie ratings are not reported over time but are merely a static measure of today's average. This implies that a single coefficient is read as "The genre X produced in the year Y is positively/negatively associated with the average movie rating of today."

| Name of Variable | Contents of Variable                                                                                                                                             | Used as .. in Analysis |
|------------|-------------------------------------------------|------------|
| Average Rating   | The average rating on IMBd per title of a given movie or series.                                                                                                 | Dependent Variable     |
| Genre Types      | Genre Types is a cluster for all 26 genre dummies that contain the value 1 if the movie or series contains that genre and otherwise 0.                           | Independent Variable   |
| titleType        | This variable indicates whether a title is a movie or a series.                                                                                                  | Fixed Effect           |
| Title            | The title is the given title of any unique movie or series                                                                                                       | Fixed Effect           |
| 5 Year Window    | The 5 year window (or ' slide_window_5' indicates between what 5 year window a movie or series belongs to. For example a movie from 2003 belonging to 2000-2004 | Split Variable          |

### Descriptives

The data set contains of 70087 films, of which 60246 are movies and 9841 are series.  These films have on average a rating of 6.3, with 10 being the highest and 1 being the lowest. After removing the votes to all being above 100, the average number of votes on a film became 7146. When looking at the years, the most amount of films were released on IMDb in 2019, with 2586 films. Also when comparing the 5 year range of films released,2015 to 2019 had the most films released, with 11353 films. LAst, when looking at the run time of the films, on average a film is 101 minutes long.

### Outcome Variable: Average Movie Rating

The average IMDb rating is used as the outcome variable for the
analysis. However, this average rating is influenced by the number of
votes that compile this average. This can be the case because each
rating, including the extreme ones, receives more weight when
calculating the average, simply because there are only a few ratings.
So, the average ratings may be more extreme, causing to skew the overall
results. Therefore, movies/series receiving a number of votes below the
cut-off of a hundred will not be included in the regression analysis.

Please note the average movie rating is a static measure that does not change the dynamics of the rating. For example, we can only see today's average rating but not that of 2 hours, 2 days, or even 2 years ago.

### Explanatory Variable: Genres

Genres are used as the explanatory variable in the analysis. However,
because they are nested within the original data set they are first
un-nested and then pivoted wider. As a result data with 27 dummies was
constructed that could be assessed separately in the analysis.

### Fixed Effects

Given that the measurement of genre preferences over time is a construct
that is viable to many unobservables, fixed effects were added to the 
regression analysis to partially capture these unobservables. More 
specifically the type of title (movie vs. series) and title were added. 
To give an example, it could be that for movies people prefer something 
with a little more drama than they would for a series or vice-versa.

A point of concern, however, is that these average ratings are not
measured over time but are a fixed sum up to and including the data of
analysis. This means that if series fall out of favor and their rating
drops this remains unobservant. Given that the continuation of a series
is determined by its success (a second season is not produced if the
first one was a flop) the following assumption can be made: if a genre
is not preferred throughout the continuance of a series, the rating is
low in the starting year and thus unlikely to be significant.
Additionally, since IMDb was not around in 1925, average ratings for
movies and series originating from that time may be strongly biased
because a niche selection of the population may be watching these movies
in the first place. Therefore, regardless of the additions of these fixed effects the results of the analysis are correlational at best.

### Capturing preferences over time

Last, to see how preferences change over time we build on the assumption
that if a movie or series is a success they gain many positive ratings
in the beginning and less so over time. As a result, there is a peak
during (the first season) release allowing us to only use the startYear
as a means to calculate behavior over time. Yet, the individual years
are unlikely to have a remarkable effect on the genre preference trends.
This is because the effects are so nuanced that they are unlikely to be
detected. Therefore, a five-year window was created that captures all 
genre preferences for that period of time. This window was constructed by grouping all productions in an interval of five years.

### Assumption

All assumptions associated with regression were checked via a visual
inspection. Only normality appeared to be satisfied.

#### Normality

```{r echo=FALSE, warning=FALSE}
library(knitr)

knitr::include_graphics("../../gen/data_preparation/qq_normality.png")
```

Homoscedasticity and independence appeared to be moderate at best.

#### Homoskedacity

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/homoskedacity.png")
```

#### Independence

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/independence.png")
```

#### Linearity

Linearity appeared to be violated the most with a widening trend for all
genres as the years evolved.

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/linearity.png")
```

Regardless of these results the decision was made to continue with the
analysis. This is because the R package `fixest` allows for modification
in the mixed model analysis that helps overcome these limitations. For
example, the package allows for clustering of different groups of
residuals. While this does not entirely mitigate the assumption concerns
it does help with them.

## Results and Discussion

The results indicated that there are different genre preferences over
time. Out of all twenty-five genres, only eighteen appeared to
significantly influence the average ratings on IMDb. These genres were
animation, biography, film-noir, fantasy, documentary, crime, comedy,
game-show, history, horror, musical, shorts, sci-fi, romance,
reality-tv, news, and war.

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_analysis/results_plot.png")
```

The genre history fell in favor in around the year 1960 and continued to
significantly (p~1960-1964~\<.01) relate to the average rating until
this day (p~2020-2024~ \<.001). While the genre film-noir stoped to
significantly relate to the average ratings around the year 1950
(p~1950-1954~=.06 , p~1955-1959~ =.49). The genres crime (p~2000-2004~
\<.01,p~2005-2009~ \<.01), romance (p~1940-1944~ =.03, p~1945-1949~
\<.01), horror (p~1925-1929~ =.04), game-show (p~2000-2004~ \<.01),
reality-tv (p~1970-1974~ \<.01), sci-fi (p~1935-1939~ \<.01), and
shorts (p~1995-1999~ \<.01, p~2005-2009~ \<.01) had a single period
bursts where they significantly related to the average ratings. Last, the
genres animation (p\~1940-1944 \<.01, p~1950-1954~ \<.01, p~1975-1979~
\<.01 -up until 2024), biography (p~1935-1939~ \<.01, p~1985-1989~ \<.01
-Up until 2024), comedy (p~1930-1934~ =.02, p~1950-1954~ =.04,
p~1990-1994~ =.01), documentary (p~1930-1934~ \<.01, p\~1945-1949 \<.01,
p\~1980-1984 \<.01 - up to and including p\~2015-2019 \<.01), fantasy
(p~1930-1934~ \<.01, p~1935-1939~ \<.01, p~1945-1949~ \<.01,
p~1970-1974~ =.03), musical (p~1960-1964~ =.09, p~1985-1989~ =.03,
p~1990-1994~ =.02, p~2020-2024~\<.01), and war (p~1930-1934~ =.02,
p~1945-1949~ \<.01, p~1960-1964~ =.04 up to and including p~1970-1974~
\<.01, p~1985-1989~ \<.01 up to and including p~2005-2009~ \<.01)
were significantly related to the average ratings in bursts with unrelated
periods in between.

A reason why genres such as animation, biography, and history are so
often correlated with the average rating over time is because these
genres hold very unique elements. For example, take romance. In romance,
the plays and sequence of events are often set and fail to vary. A
watcher often knows what is going to happen next and what it may look
like. Controversially, the genres of animation, biography, and history
contain more unique and unpredictable elements. For example, not every
animated movie is Frozen or the Incredibles, the history of Egypt is not
the same as that of Amelia Earhart or even Germany, and the story of
Elton John is different from that of Justin Timberlake.

For genres like war and film-noir the correlation could be a sign of the
times, and can be strongly guided by biases in the rating population. For
example, when the genre war was indicated to be significantly related to
average ratings it was the time the Second World War and the Cold War
were in full motion. Therefore people who have an interest in this topic
may like to watch propaganda movies from that time. As a result, the
average ratings are higher because a niche of interested people is
watching. Similar to film-noir, people who are interested in this type
of movie or series are watching these and rating them. Therefore, there
may be a strong instance of viewer bias present.

On the contrary, the genres action, adult, adventure, mystery, sport,
thriller, and western are all insignificant and do not appear to be
correlated to the average ratings on IMDb. A reason for this may be that
these elements have become expected or core components of movies and
series and therefore do not stand out. As a result, they may not
correlate with the average rating as much as elements of a biography may
do. It is not every day a movie or series is released that is based on
true events.

# Conclusion

This research illustrates that genre preferences may be related to the
average IMDb ratings and vary over time. While further research is
necessary to fine-tune these claims the current research does provide
insight into how raters' watch patterns appear to alter. For example, it
shows that the common ingredients like romance and action may not stand
out as much as biography and history might.

The results of this analysis can be, at best, interpreted as descriptive 
given the vast amount of assumptions and potential hidden dangers of 
sampling biases (e.g., film-noir and war genres). While this has been 
partially mitigated by loosening some of the underlying assumptions (clustering standard errors and selecting a better heterogeneity method) and adding fixed effects, the fact remains that these appear as trend 
observations. Therefore, the results do not necessarily conclude that some 
genres lead to higher average ratings. However, the results do indicate that there may be preferences over time that may lead to higher average ratings.

## Extensions

Yet, the investigation into these genres is only a starting point. Much
more can be learned and investigated. Here are three ways in which the
current research into how genres relate to average IMDb rating may be
extended:

First, researchers may want to make use of dynamic review data that is
not an accumulation of a point in time. Currently, the sole measure of
the average reviews resulted in two major assumptions in the analysis.
The first assumption refers to series and franchise continuation. E.g.,
if the first film or series was not successful there is no follow-up.
And the second assumption refers to the fact that the year window is
based on the startYear of the movie and series. By looking at dynamic
data researchers can also observe if genres fall out of favor in
specific conditions. For example, as the series continues do people tire
of the interpretation of the genre, or is the genre in general less
popular?

Second, researchers may benefit by including data from multiple sources.
Currently, the analysis is limited to IMDb data but there are also other
sources such as Rotten Tomatoes for example. By combining these data sets
the interpretation of the results becomes less biased by the data
platform. For example, it may be that people of higher age groups
typically review on IMDb, and younger groups adopt commentators. By
broadening the data sources the generalization of the results increases.

Last, researchers may benefit by including additional control variables
such as GDP and reported population happiness scales. Given that the
movie and series market is an indulgent market it may be that genres did
not fall out of favor but that people could not afford to watch movies
and series. Or it may be that other matters were stirring in the
population which made them more happy or sad. These factors influence
consumption patterns and thus ratings. By broadening the scope of the
analysis results become more reliable.